\section{Problem Statement}

In this project, we generate explanations for predictions made by a deep neural network that has been trained on the Yelp dataset \cite{Yelp2017}. We generate explanations for two prediction tasks: 1) Sentiment analysis, and 2) Gender prediction.

\subsection{Sentiment Analysis}

Yelp reviews are utilized by users to make decisions about which service provider (restaurants, doctors, etc.) to use for their needs. In making these decisions, users need to understand what aspects of the service provided is reflected in a reviewer's overall rating. For example, if a user is interested in a restaurant which has low ratings, they might need to understand if the low rating is because of bad food or some other aspect (such as ambience, etc.) which they do not care about. So, it is essential to enumerate the factors that contributed to a given reviewer's rating.

\subsection{Gender Prediction}

Recent advances in natural language processing have enabled the prediction of an author's gender from the text they have written by automatic analysis of linguistic characteristics of the text. We investigate the accuracy of deep neural networks in predicting the gender of a user based on the reviews that they have written. Again, it is important to understand which factors of the review text contribute towards making an accurate prediction. This would provide an understanding of what needs to be done to change the review text suitably in case a user wants to provide an anonymous review.
