\section{Related Work}

A recurrent neural network (RNN) is a neural network which allows cyclical connections. Thus, recurrent connections allow a `memory' of previous inputs to persist in the network's internal state, which can then be used to influence the network output. Many varieties of RNNs have been proposed in the literature, such as Elman networks \cite{Elman1990}, Jordan networks \cite{Jordan1990}, time delay neural networks \cite{Lang1990}, etc.

In standard RNN architectures, the influence of a given input on the network output either decays or blows up exponentially as it cycles around the network's recurrent connections. This shortcoming is referred to as the \textit{vanishing gradient problem} \cite{Bengio1994}. The most effective solution so far to the vanishing gradient problem is the Long Short Term Memory (LSTM) architecture \cite{Hochreiter1997}.

Convolutional neural networks (CNNs) were developed to improve the effectiveness of pattern detection tasks in computer vision. They were inspired by the concept of receptive field \cite{Fukushima1988} in biological vision. The biological function can be approximated in computers using the convolution operation \cite{Marr1980}. A set of convolutional filters can be combined to form a \textit{convolutional layer} of a neural network.

In recent years, RNNs and CNNs have been used to perform a variety of natural language processing tasks. For example, in sentiment analysis tasks, most state-of-the-art methods \cite{Tang2015} use a combination of convolutional and recurrent neural networks. Convolutional neural networks, despite their use mainly in image-based applications, are also used in text-based applications which involve summarization of text since convolution can learn a holistic representation of a group of words instead of just individual words. Similar architectures of neural network models have been used in gender classification tasks as well \cite{Burger2011, Mukherjee2010}. However, none of these works generate explanations about why the model works well, or what kind of information is learnt by the system.
